---
title: "Data Handling"
output: html_notebook
Author : Swapna
---


```{r}
library(tidyverse)
```

```{r}
flight_data = read.csv("C:\\Users\\Swapna\\Desktop\\Rstudio\\Review\\nycflights.csv")
```

Take a quick peek at your data to understand its contents better.


```{r}
glimpse(flight_data)
```
# Departure delays

Examine the distribution of departure delays of all flights with a histogram.

```{r}
delay = flight_data$dep_delay
hist(delay)
```
Visualize only on delays of flights headed to Los Angeles, you need to first filter the data for flights with that destination (dest == "LAX") and then make a histogram of the departure delays of only those flights.

```{r}
flight_data%>%
  filter(dest == "LAX")%>%
  group_by("Los Angeles")%>%
  select(dest,dep_delay)
```


```{r}
library(ggplot2)

```


```{r}
ggplot(flight_data, aes(x = dep_delay , color = "LAX"))+geom_histogram()
```
summarise delays by showing the mean delay duration (mean_dd), median delay duration (median_dd), and number of delayed flights (n)

```{r}
glimpse(flight_data)
```


```{r}
flight_data%>%
  summarize(mean_dd = mean(dep_time), median_dd = median(dep_time), n = n())
```

Create a new data frame that includes flights headed to SFO in February, and save this data frame as sfo_feb_flights. How many flights meet these criteria?

```{r}
sfo_feb_flights = flight_data%>%
  filter(dest == "SFO", month == 2)
```

```{r}
sfo_feb_flights
```

Calculate the median and interquartile range for arr_delays of flights in in the sfo_feb_flights data frame, grouped by carrier. Which carrier has the most variable arrival delays?
```{r}
sfo_feb_flights%>%group_by(carrier)%>%summarise(median_arr=median(arr_delay), iqr=IQR(arr_delay))

```

Suppose you really dislike departure delays and you want to schedule your travel in a month that minimizes your potential departure delay leaving NYC. Calculate monthly averages for departure delays, then arrange these average dealys in descending order

```{r}
flight_data%>%group_by(month)%>%summarise(mean_dep_delay = mean(dep_delay))%>%arrange(desc(mean_dep_delay))
```

Mutate the data frame so that it includes a new variable that contains the average speed, avg_speed traveled by the plane for each flight (in mph). Hint: Average speed can be calculated as distance divided by number of hours of travel, and note that air_time is given in minutes.
Make a scatterplot of avg_speed vs. distance. Describe the relationship between average speed and distance. Hint: Use geom_point().


```{r}
flight_data%>%mutate(avg_speed = distance / air_time/60)%>%select(avg_speed,flight)
ggplot(flight_data, aes(x = distance , y = distance/air_time/60)) + geom_point()


```


```{r}
```






















